---
title: "aquaculture_development"
format: html
Excute:
  warning: false
---

# Homework 4:

Prioritizing potential aquaculture

Loading Libraries

```{r}
#| output: FALSE

library(tidyverse)
library(terra)
library(stars)
library(sf)
library(tmap)
library(kableExtra)
```

Downloading data

```{r}
# Sea Surface Temperature
rast_2008 <- rast(here::here('data', 'average_annual_sst_2008.tif'))
rast_2009 <- rast(here::here('data', 'average_annual_sst_2009.tif'))
rast_2010 <- rast(here::here('data', 'average_annual_sst_2010.tif'))
rast_2011 <- rast(here::here('data', 'average_annual_sst_2011.tif'))
rast_2012 <- rast(here::here('data', 'average_annual_sst_2012.tif'))

# Depth 
bathy <- rast(here::here('data', 'depth.tif'))

# Exclusive Economic Zones
eez <- read_sf(here::here('data', 'wc_regions_clean.shp'))

```

## Coordinate Reference System

Check if the CRS match and update if not

```{r}

# CRS UPDATING function 
crs_update <- function (df1, df2) {
  if (st_crs(df1) == st_crs(df2)) {
    print("CRS match")
  } else {
    warning("Updating CRS") 
    df1 <- st_transform(df2, crs = "EPSG:4326")
    df2 <- st_transform(df2, crs = "EPSG:4326") 
  }
}

# CRS CHECK function (with warning)
crs_check <- function (df1, df2) {
  if (st_crs(df1) == st_crs(df2)) {
    print("CRS match")
  } else {
    warning("CRS do NOT match") 
  }
}

# Check the dataframes
crs_update(bathy, eez)
crs_update(bathy, rast_2008)
crs_update(rast_2008, rast_2009)
crs_update(rast_2009, rast_2010)
crs_update(rast_2010, rast_2011)
crs_update(rast_2011, rast_2012)
```

## Process data

**Combine** sst rasters

```{r}
# Combine Sea Surface Temperature Rasters  
sst_rast <- c(rast_2008, rast_2009, rast_2010, rast_2011, rast_2012)
# rast() would also work to merge 

# Remove unbinded rasters 
rm(list = c('rast_2008', 'rast_2009', 'rast_2010', 'rast_2011', 'rast_2012'))
```

Find the **mean** SST - LOCAL algebra

```{r}
# Find the mean SST
sst_mean <- mean(sst_rast)
```

**Convert** SST from Kelvin to Celsius- GLOBAL algebra

```{r}
# Subtract 273.15 
sst_mean <- sst_mean - 273.15
```

**Crop** depth raster to match the *extent* of the SST raster

```{r}
# Extent of sst raster 
sst_ext <- ext(sst_mean)

# Crop bathy to sst raster extent
bathy_crop <- crop(bathy, sst_ext)

rm(sst_ext)

```

Fixing unmatching **resolutions**

```{r}
# Resample bathy / depth data to match sst raster 
bathy_updated <- resample(bathy_crop, sst_mean, method = "bilinear")

# Check resolutions 
if (all(res(sst_rast) == res(bathy_updated))) {
  print("Resolutions match")
} else {
  warning("Resolutions don't match!")
}

# Remove uncropped and unresolutioned depth dfs 
rm(list = c('bathy', 'bathy_crop'))
```

Final Check depth and SST dfs match in resolution, extent, and coordinate reference system

```{r}
# Do resolutions match? 
all(res(bathy_updated) == res(sst_mean)) # Equal in both vs all directions
# Do extents match?
ext(bathy_updated) == ext(sst_mean) 
# Do CRS match? 
crs_check(bathy_updated, sst_mean)

```

------------------------------------------------------------------------

## Find suitable locations for [Oysters]{.underline}

Oysters Research has shown that oysters need the following conditions for optimal growth:

-   Sea surface temperature: 11-30°C
-   Depth: 0-70 meters below sea level

Visualize Depth vs SST of West Coast

```{r}
#| code-fold: true

# Map sst and bathy 
bath_map <- tm_shape(bathy_updated) + 
  tm_raster(col.legend = tm_legend(title = "Depth (m)")) + 
  tm_title(text = "Depth (m) of West Coast") + 
  tm_compass(position = c('right', 'top'))

sst_map <- tm_shape(sst_mean) + 
  tm_raster(col.legend = tm_legend(title = "Temperature (°C)"), 
            col.scale = tm_scale(values = c('#ffcc84',"#FFB627", "#ff8c34", "#CC5803", "#772f1a"))) + 
  tm_title(text = "Average Sea Surface Temperature\nof West Coast") + 
  tm_scalebar(position = c('right', 'top')) + 
  tm_basemap("Esri.OceanBasemap")

tmap_arrange(sst_map, bath_map)

```

### Reclassifying locations that have optimal oyster temp or depth

-   Unsuitable habitats = 0
-   Suitable habitats = 1

Suitable **DEPTH** (0-70 meters below sea level) vs Unsuitable locations for oysters

```{r}
# Create a reclassification matrix of values corresponding to unsuitable (1) vs suitable (2) depths for oysters 
rcl_depth <- matrix(c(-Inf, -70, 0, # SUITABLE depth
                      -70, 0, 1, #  SUITABLE depth
                      0, Inf, 0), # SUITABLE depth
                      ncol = 3, byrow = TRUE)

# Apply the reclassification matrix to the raster 
depth_reclass <- classify(bathy_updated, rcl_depth) #other = NA)

# Delete matrix as object 
rm(rcl_depth)
```

Suitable **TEMP** (11-30°C) vs Unsuitable locations for oysters

```{r}
# Create a reclassification matrix of values corresponding to unsuitable (1) vs suitable (2) temps for oysters 
rcl_temp <- matrix(c(-Inf, 11, 0, # UNSUITABLE temp 
                     11, 20, 1, #SUITABLE temp
                     20, Inf, 0), #UNSUITABLE temp
                   ncol = 3, byrow = TRUE)

# Apply the reclassification matrix to the raster 
temp_reclass <- classify(sst_mean, rcl_temp) #other =NA)

# Delete matrix as object 
rm(rcl_temp)
```

### Finding which areas have suitable depth *AND* temp for Oysters

```{r}
#| warning: false

# Check CRSs
crs_check(temp_reclass, depth_reclass)


# Create multiplication function
multiply <- function(layer1, layer2) { 
  layer1 * layer2
}

# Apply function to rasters 
oyst_suit_raster <- lapp(c(temp_reclass,depth_reclass), fun = multiply)

```

Visualizing depth and temp suitability for oysters

```{r}
#| code-fold: true
#| warning: false

# Plotting suitable habitats for Oysters 
map_temp <- tm_shape(temp_reclass) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1) [11-30°C]"), 
      values = c('lightblue2', 'steelblue4'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Temp"))+ 
  tm_title(text = "Suitable Temp for Oysters") + 
  tm_compass(position = c('right', 'top'))


map_depth <- tm_shape(depth_reclass) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1) [0-70 meters]"), 
      values = c('lightblue2', 'steelblue4'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Depth")
  ) + 
  tm_title(text = "Suitable Depth for Oysters")


map_oyst <- tm_shape(oyst_suit_raster) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1)"), 
      values = c('lightblue2', 'red'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Habitat")) + 
  tm_title(text = "Suitable Depth & Temp for Oysters") + 
  tm_scalebar(position = c('left', 'bottom')) 

tmap_arrange(map_temp, map_depth, map_oyst)

# Remove maps as objects 
rm(list = c('map_temp', 'map_depth', 'map_oyst'))

```

### Determining Most Suitable EEZ

Visualizing EEZs and suitable oysters habitat with EEZs

```{r}
#| code-fold: true
#| warning: false

# Vizualization of Habiatat and EEZs 
tm_shape(oyst_suit_raster) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1)"), 
      values = c('lightblue2', 'red'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Habitat")) + 
  tm_title(text = "Suitable Depth & Temp for Oysters with\nthe West Coast Exclusive Economic Zones") + 
  tm_shape(eez) + 
  tm_polygons( 
              lwd = 0.35, 
              fill = "peachpuff2", 
              fill_alpha = 0.2) + 
  tm_scalebar(position = c('left', 'bottom'), 
              margins = 1
              ) +
  tm_compass(position = c('right', 'top')) + 
  tm_credits(text = "Oysters perfer depths of 0-70m and temps of 11-30°C.")

```

Selecting Suitable Habitats within all the West Coast EEZs

```{r}
#| warning: false

# Check is crs match 
crs_check(oyst_suit_raster, eez)

# Mask suitable hab raster to the EEZ area 
oyst_suit_crop <- mask(oyst_suit_raster, eez)
```

Check if masked correctly by plotting

```{r}
#| code-fold: true

# Check if it masked 
tm_shape(oyst_suit_crop) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1)"), 
      values = c('lightblue2', 'red'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Habitat")) + 
  tm_title(text = "Suitable Depth & Temp for Oysters within\nthe West Coast Exclusive Economic Zones") + 
  tm_shape(eez) + 
  tm_polygons( 
              lwd = 0.25, 
              fill = "peachpuff2", 
              fill_alpha = 0.2) + 
  tm_scalebar(position = c('left', 'bottom'), 
              margins = 1
              ) +
  tm_compass(position = c('right', 'top')) + 
  tm_credits(text = "Oysters perfer depths of 0-70m and temps of 11-30°C.") + 
  tm_basemap("Esri.OceanBasemap")

```

Find **Area** of Suitable habitat in West Coast EEZs

Total Area of West Coast EEZ

```{r}
# Total area of all West Coast EEZ 
print(paste("The total area of all West Coast EEZs is", round(sum(st_area(eez))/1e9,2), "x 10^9 m^2"))

```

Total Oyster suitable habitat in West Coast EEZ

```{r}
# Creating raster of JUST the suitable habitats [1]
oyst_1 <- oyst_suit_crop
oyst_1[oyst_1  == 0] <- NA # unsuitable hab are 'NAs'


# Find total area of suitable habitats 
area <- expanse(oyst_1)[1,2] 
# Expense computes the area covered by raster cells that are not NA 

print(paste("The area of all suitable oyster habitats in the West Coast EEZs is", round(area/1e9, 2), "x 10^9 m^2"))

```

Create a generalized function to find the area of suitable habitat for a species per each West Coast EEZ

```{r}
# Find area of suitable habitat in each EEZ 
suit_area_eez_generalized <- function (key, eez_zone_df, species_raster) { 
  
  # Make sure the crs match - warning if they do not match 
  if (st_crs(eez_zone_df) == st_crs(species_raster)) {
    print("CRS match")
  } else {
    warning("CRS do NOT match!")
  }
  
  
  area_eez <- eez_zone_df %>% filter(rgn == key) # Filter per EEZ 
  spec_suit <- mask(species_raster, area_eez) # Masking the cropped-habitat-raster with each EEZ location
  area <- expanse(spec_suit)[1,2] # Find area of the cropped hab rast 
 
  area_format <-  paste(round(area / 1e8, 2), "x 10^8") # Make area easier to comprehend 
  
  #print(paste(key, "has an area of", area, "of oyster suitable habitat"))
  #print(paste(round(area/10^9, 2), "x 10^9"))
  df <- data.frame(key,area = area_format) # Make a dateframe out of results 
  return(df)
  
}
```

Apply the function to Oyster data

```{r}
#| warning: false

# List of all eez keys 
eez_keys <- c('Washington', 'Oregon', 'Northern California', 'Central California', 'Southern California') 


# Apply the function to all keys and Oyster dataframe and make a combined dataframe from it 
df_suit_area_eez  <- do.call( 
  rbind, # Combine outputs into 1 data frame 
  lapply( # Apply all Oyster data to the function 
    eez_keys,
    suit_area_eez_generalized,
    eez_zone_df = eez,
    species_raster = oyst_1
  )
)


# Create a Table 
kable(df_suit_area_eez, align = 'c', col.names = c("EEZ location", "Oyster Suitable Habitat Area"))

```

Oysters have a large amount of suitable habitats in Washington, Central CA, and Southern CA. Central and South California's temperature, ranging from 12°C to 20°C, is fully within the limits of oysters preferences. Additionally, oysters have a depth span of 0-70 meters below sea level, meaning oysters will reside closer to the coastline. The Washington and South California's Exclusive Economic Zones offer a larger continental shelf where oysters can thrive. In conclusion, when mixing oysters depth and temperature preferences, Washington, Central CA, and Southern CA offer the most suitable habitats.

# Suitable locations for [Siliqua-patula]{.underline}

The Pacific razor clam location ranges from the Aleutian Islands in Alaska to Pismo Beach, California. The clam is typically discovered in depths between 0-55m, and its preferred temperature is around 5°C (estimating 0°C - 10°C).

Therefore, Pacific razor clam research has shown that the clams need the following conditions for optimal growth:

-   Sea surface temperature: 0-10°C

-   Depth: 0-55 meters below sea level

https://www.sealifebase.ca/summary/Siliqua-patula.html

### Reclassifying locations that have optimal clam temp or depth

Creating a suitable habitat **matrix** for the Pacific razor clam

```{r}
# Depth matrix 
rcl_depth_clam <- matrix(c(-Inf, -55, 0, # UNSUITABLE
                      -55, 0, 1, #  SUITABLE HABITAT
                      0, Inf, 0), # UNSUITABL
                      ncol = 3, byrow = TRUE)
# Temp matrix 
rcl_temp_clam <- matrix(c(-Inf, 0, 0, # UNSUITABLE
                      0, 10, 1, #  SUITABLE HABITAT
                      10, Inf, 0), # UNSUITABL
                      ncol = 3, byrow = TRUE)

# Apply the reclassification matrix to the depth and temp rasters 
depth_reclass_clam <- classify(bathy_updated, rcl_depth_clam)

temp_reclass_clam <- classify(sst_mean, rcl_temp_clam)

# Remove matrixs as objects 
rm(list = c('rcl_depth_clam', 'rcl_temp_clam'))

```

### Finding which areas have suitable depth *AND* temp for Clams

Multiplying rasters together - Suitable habitat for Pacific razor clam

```{r}
#| warning: false

# Check is crs match 
crs_check(depth_reclass_clam, temp_reclass_clam)

# Multiple temp and depth rasters 
oyst_suit_raster <- lapp(c(depth_reclass_clam, temp_reclass_clam), fun = multiply)

```
### Determining Most Suitable EEZ

Masking the Suitable habitat for Pacific razor clam raster to the West Coast EEZ

```{r}
clam_suit_crop <- mask(oyst_suit_raster, eez)
```

Visualize everything
```{r}
#| code-fold: true

# Plotting suitable habitats for Oysters 

# TEMP 
map_temp_clam <- tm_shape(temp_reclass_clam) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1) [0-10°C]"), 
      values = c('lightblue2', 'steelblue4'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Temp"))+ 
  tm_title(text = "Suitable Temp for Clams") + 
  tm_compass(position = c('right', 'top'))

# DEPTH 
map_depth_clam <- tm_shape(depth_reclass_clam) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1) [0-55 meters]"), 
      values = c('lightblue2', 'steelblue4'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nClams Depth")
  ) + 
  tm_title(text = "Suitable Depth for Clams")

# TEMP AND DEPTH 
map_clam <- tm_shape(oyst_suit_raster) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1)"), 
      values = c('lightblue2', 'red'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nClams Habitat")) + 
  tm_title(text = "Suitable Depth & Temp for Clams") + 
  tm_scalebar(position = c('left', 'bottom')) 

map_clam_crop <- tm_shape(clam_suit_crop) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1)"), 
      values = c('lightblue2', 'red'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Habitat")) + 
  tm_title(text = "Suitable Depth & Temp for Clams within the West Coast EEZs") + 
  tm_shape(eez) + 
  tm_polygons(lwd = 0.35, 
              fill = "peachpuff2", 
              fill_alpha = 0.2) + 
  tm_basemap("CartoDB.PositronNoLabels")



tmap_arrange(map_temp_clam, map_depth_clam, map_clam)

```

```{r}
#| code-fold: true
#| warning: false

# TEMP AND DEPTH AND EEZ 

# bbox of Washington EEZ 
washington <- eez %>% filter(rgn_key == 'WA')
clam_suit_washington <- mask(clam_suit_crop, washington)


# Inset map of Just clams in Washington 
washington_inset <- tm_shape(
   clam_suit_washington, 
   bbox = sf::st_bbox(c(
    xmin = -129.5, xmax = -122,
    ymin = 45.5,  ymax = 49
  )), 
  crs = st_crs(clam_suit_washington) ) + 
  tm_raster( 
    col.scale = tm_scale_categorical(
      values = c('lightblue2', 'red')), 
    col.legend = tm_legend_hide()) +
  tm_shape(washington) + 
  tm_polygons(lwd = 0.35, 
              fill = "peachpuff2", 
              fill_alpha = 0.2) +
  tm_title(text = "Washington's EEZ", 
           position = tm_pos_in(), 
           size = 0.8) +
 tm_basemap("Esri.OceanBasemap") + 
  tm_scalebar(position = c('right', 'bottom'))

 
# TEMP AND DEPTH AND EEZ MAP 
tm_shape(
  clam_suit_crop, 
     bbox = sf::st_bbox(c(
    xmin = -131, xmax = -110,
    ymin = 30,  ymax = 50
  )), 
  crs = st_crs(clam_suit_crop)) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1)"), 
      values = c('lightblue2', 'red'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Habitat", 
                           position = tm_pos_in('right', 'center'), 
                           title.size = 0.6, 
                           text.size = 0.5, 
                           bg.alpha = 0.5)
    ) + 
  tm_title(text = "Suitable Depth & Temp for Clams\nwithin the West Coast EEZs") + 
  tm_shape(eez) + 
  tm_polygons(lwd = 0.35, 
              fill = "peachpuff2", 
              fill_alpha = 0.2) + 
  tm_basemap("Esri.OceanBasemap") +
  tm_compass(position = c('left', 'bottom'))+ 
  tm_scalebar(position = c('left', 'bottom')) +
  tm_inset(washington_inset, 
           position = tm_pos_out(), 
           width = 15, 
           height = 15)



# Remove objects 
rm(list = c('map_temp_clam', 'map_depth_clam', 'map_clam'))

```


Total Area of Clam suitable habitat in all West Coast EEZs

```{r}
#| warning: false 
clam_1 <- clam_suit_crop
clam_1[clam_1   == 0] <- NA # unsuitable hab are 'NAs'

# Find total area of suitable habitats 
area <- expanse(clam_1)[1,2] 
# Expense computes the area covered by raster cells that are not NA 

print(paste("The area of all suitable Pacific razor clam habitats in the West Coast EEZs is", round(area/1e8, 2), "x 10^8 m^2"))
```

Which EEZ has the most clam suitable habitat

```{r}
# Apply the function to all keys and Oyster dataframe and make a combined dataframe from it 
df_suit_area_eez  <- do.call( 
  rbind, # Combine outputs into 1 data frame 
  lapply( # Apply all Oyster data to the function 
    eez_keys,
    suit_area_eez_generalized,
    eez_zone_df = eez,
    species_raster = clam_1
  )
)


# Create a Table 
kable(df_suit_area_eez, align = 'c', col.names = c("EEZ location", "Pacific Razor Clam Suitable Habitat Area in West Coast EEZs"))


```



The Pacific razor clams are mostly found on the continental shelf with shallow depths from 0 to 55 meters - which is mostly offered in the North. Furthermore, the clams need a temperature of around 5°C which limits them to the northern parts of the West Coast as well. Hence, Washington has the largest area of suitable habitat for the Pacific Razor Clams. 

