---
title: "EEZ Aqauculture Development"
author: Megan Hessel 
date: November 24, 2025
execute:
  warning: false
  message: false
format:
  pdf: default
  html: default 
---

# Prioritizing Potential Aquaculture

This repository investigates temperature and depth requirements for different aquaculture species on the West Coast.

Loading Libraries

```{r}
#| output: FALSE

library(tidyverse)
library(terra)
library(stars)
library(sf)
library(tmap)
library(kableExtra)
```

Downloading data - Temperature, depth, and West Coast EEZ data

```{r}
# Sea Surface Temperature
rast_2008 <- rast(here::here('data', 'average_annual_sst_2008.tif'))
rast_2009 <- rast(here::here('data', 'average_annual_sst_2009.tif'))
rast_2010 <- rast(here::here('data', 'average_annual_sst_2010.tif'))
rast_2011 <- rast(here::here('data', 'average_annual_sst_2011.tif'))
rast_2012 <- rast(here::here('data', 'average_annual_sst_2012.tif')) 

# Depth 
bathy <- rast(here::here('data', 'depth.tif')) %>% 
  project('EPSG:4326')

# Exclusive Economic Zones
eez <- read_sf(here::here('data', 'wc_regions_clean.shp')) 

```

## Coordinate Reference System

Check if the CRS match and update if not

```{r}

# CRS UPDATING function 
crs_update <- function (df1, df2) {
  if (st_crs(df1) == st_crs(df2)) {
    print("CRS match")
  } else {
    warning("Updating CRS") 
    df1 <- st_transform(df2, crs = "EPSG:4326")
    df2 <- st_transform(df2, crs = "EPSG:4326") 
  }
}

# CRS CHECK function (with warning)
crs_check <- function (df1, df2) {
  if (st_crs(df1) == st_crs(df2)) {
    print("CRS match")
  } else {
    warning("CRS do NOT match") 
  }
}

# Check the dataframes
crs_update(bathy, eez)
crs_update(bathy, rast_2008)
crs_update(rast_2008, rast_2009)
crs_update(rast_2009, rast_2010)
crs_update(rast_2010, rast_2011)
crs_update(rast_2011, rast_2012)

```

## Process data

**Combine** sst rasters

```{r}
# Combine Sea Surface Temperature Rasters  
sst_rast <- c(rast_2008, rast_2009, rast_2010, rast_2011, rast_2012)
# rast() would also work to merge 

# Remove unbinded rasters 
rm(list = c('rast_2008', 'rast_2009', 'rast_2010', 'rast_2011', 'rast_2012'))
```

Find the **mean** SST - local algebra

```{r}
# Find the mean SST
sst_mean <- mean(sst_rast)
```

**Convert** SST from Kelvin to Celsius - global algebra

```{r}
# Subtract 273.15 
sst_mean <- sst_mean - 273.15
```

**Crop** depth raster to match the *extent* of the SST raster

```{r}
# Extent of sst raster 
sst_ext <- ext(sst_mean)

# Crop bathy to sst raster extent
bathy_crop <- crop(bathy, sst_ext)

rm(sst_ext)

```

Fixing unmatching **resolutions**

```{r}
# Resample bathy / depth data to match sst raster 
bathy_updated <- resample(bathy_crop, sst_mean, method = "bilinear")

# Check resolutions 
if (all(res(sst_rast) == res(bathy_updated))) {
  print("Resolutions match")
} else {
  warning("Resolutions don't match!")
}

# Remove uncropped and unresolutioned depth dfs 
rm(list = c('bathy', 'bathy_crop'))
```

Final Check depth and SST dfs match in resolution, extent, and coordinate reference system

```{r}

# Do resolutions match? 
all(res(bathy_updated) == res(sst_mean)) # Equal in both vs all directions
# Do extents match?
ext(bathy_updated) == ext(sst_mean) 
# Do CRS match? 
crs_check(bathy_updated, sst_mean)

```

## Temp and Depth Map of West Coast

```{r}
#| code-fold: true

# Map sst and bathy 
bath_map <- tm_shape(bathy_updated) + 
  tm_raster(col.legend = tm_legend(title = "Depth (m)")) + 
  tm_title(text = "Depth (m) of West Coast") + 
  tm_compass(position = c('right', 'top'))

sst_map <- tm_shape(sst_mean) + 
  tm_raster(col.legend = tm_legend(title = "Temperature (°C)"), 
            col.scale = tm_scale(values = c('#ffcc84',"#FFB627", "#ff8c34", "#CC5803", "#772f1a"))) + 
  tm_title(text = "Average Sea Surface\nTemperature of West Coast") + 
  tm_scalebar(position = c('right', 'top')) + 
  tm_basemap("Esri.OceanBasemap")

tmap_arrange(sst_map, bath_map)

```

## Find suitable locations for [Oysters]{.underline}

Oysters Research has shown that oysters need the following conditions for optimal growth:

-   Sea surface temperature: 11-30°C
-   Depth: 0-70 meters below sea level

### Reclassifying locations that have optimal oyster temp or depth

-   Unsuitable habitats = 0
-   Suitable habitats = 1

Suitable **DEPTH** (0-70 meters below sea level) vs Unsuitable locations for oysters

```{r}
# Create a reclassification matrix of values corresponding to unsuitable (1) vs suitable (2) depths for oysters 
rcl_depth <- matrix(c(-Inf, -70, 0, # SUITABLE depth
                      -70, 0, 1, #  SUITABLE depth
                      0, Inf, 0), # SUITABLE depth
                      ncol = 3, byrow = TRUE)

# Apply the reclassification matrix to the raster 
depth_reclass <- classify(bathy_updated, rcl_depth) #other = NA)

# Delete matrix as object 
rm(rcl_depth)
```

Suitable **TEMP** (11-30°C) vs Unsuitable locations for oysters

```{r}
# Create a reclassification matrix of values corresponding to unsuitable (1) vs suitable (2) temps for oysters 
rcl_temp <- matrix(c(-Inf, 11, 0, # UNSUITABLE temp 
                     11, 30, 1, #SUITABLE temp
                     30, Inf, 0), #UNSUITABLE temp
                   ncol = 3, byrow = TRUE)

# Apply the reclassification matrix to the raster 
temp_reclass <- classify(sst_mean, rcl_temp) #other =NA)

# Delete matrix as object 
rm(rcl_temp)
```

### Finding which areas have suitable depth *AND* temp for Oysters

Multiplying temp and depth raster together

```{r}

# Check CRSs
crs_check(temp_reclass, depth_reclass)

# Multiply rasters together 
oyst_suit_raster <- temp_reclass * depth_reclass

# OTHER METHOD: FUNCTION WITH LAPP 
# multiply <- function(layer1, layer2) { 
#   layer1 * layer2
# }
# oyst_suit_raster <- lapp(c(temp_reclass,depth_reclass), fun = multiply)

```

Visualizing depth and temp suitability for oysters in the west coast

```{r}
#| code-fold: true
#| warning: false

# Plotting suitable habitats for Oysters 
map_temp <- tm_shape(temp_reclass) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1) [11-30°C]"), 
      values = c('lightblue2', 'steelblue4'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Temp"))+ 
  tm_title(text = "Suitable Temp for Oysters") + 
  tm_compass(position = c('right', 'top'))


map_depth <- tm_shape(depth_reclass) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1) [0-70 meters]"), 
      values = c('lightblue2', 'steelblue4'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Depth")
  ) + 
  tm_title(text = "Suitable Depth for Oysters")


map_oyst <- tm_shape(oyst_suit_raster) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1)"), 
      values = c('lightblue2', 'red'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Habitat")) + 
  tm_title(text = "Suitable Depth & Temp for Oysters") + 
  tm_scalebar(position = c('left', 'bottom')) 

tmap_arrange(map_temp, map_depth, map_oyst)

# Remove maps as objects 
rm(list = c('map_temp', 'map_depth', 'map_oyst'))

```

## Determining most suitable EEZ

**Mask** the Oyster raster to the EEZ geometries

```{r}

# Check is crs match 
crs_check(oyst_suit_raster, eez)

# Mask suitable hab raster to the EEZ area 
oyst_suit_crop <- mask(oyst_suit_raster, eez)
```

### Plot Oyster suitability *within* the EEZ areas

```{r}
#| code-fold: true

# Check if it masked 
tm_shape(oyst_suit_crop) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1)"), 
      values = c('lightblue2', 'red'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Habitat")) + 
  tm_title(text = "Suitable Depth & Temp for Oysters within\nthe West Coast Exclusive Economic Zones") + 
  tm_shape(eez) + 
  tm_polygons( 
              lwd = 0.25, 
              fill = "peachpuff2", 
              fill_alpha = 0.2) + 
  tm_scalebar(position = c('left', 'bottom'), 
              margins = 1
              ) +
  tm_compass(position = c('right', 'top')) + 
  tm_credits(text = "Oysters perfer depths of 0-70m and temps of 11-30°C.") + 
  tm_basemap("Esri.OceanBasemap")
```

### Find **Area** of Suitable habitat in West Coast EEZs

Total Area of all West Coast EEZ

```{r}
# Total area of all West Coast EEZ 
print(paste("The total area of all West Coast EEZs is", round(sum(st_area(eez))/1e9,2), "x 10^9 m^2"))

```

Total Oyster suitable habitat in West Coast EEZ

```{r}
# Creating raster of JUST the suitable habitats [1]
oyst_1 <- oyst_suit_crop
oyst_1[oyst_1  == 0] <- NA # unsuitable hab are 'NAs'


# Find total area of suitable habitats 
area <- expanse(oyst_1)[1,2] 
# Expense computes the area covered by raster cells that are not NA 

print(paste("The area of all suitable oyster habitats in the West Coast EEZs is", round(area/1e9, 2), "x 10^9 m^2"))

```

**Trying new things**
```{r}

for (i in unique(eez$rgn)) {
  zone <- eez %>% filter(rgn == i)
  crs_check(zone, oyst_1)
  oyst_zone <- mask(oyst_1, zone)
  area <- expanse(oyst_zone)[1,2] 
  
  zone[, 'area_suit_hab'] <- NA
  zone$area_suit_hab <- area
 
  df_list <- list()
  df_list[[i]] <- data.frame(
    rgn = i, 
     area_per_eez = paste(round(area / 1e8, 2), "x 10^8")
  )
  
  }

  eez_plus <- do.call(rbind, df_list)
  eez_plus
  
  # table <- eez_plus %>% 
  #   select(rgn, area_suit_hab) %>% 
  #   st_drop_geometry()

  #table

```


Initial generalized function to find the area of suitable habitat for a species per each West Coast EEZ

```{r}

df_list <- list()
  
# For loop to get area of each eez
for (i in unique(eez$rgn)) {
    area_eez <- eez %>% filter(rgn == i) #
    spec_per_eez <- mask(oyst_1, area_eez)
    area_per_eez <- expanse(spec_per_eez)[1, 2]
    
    # Create dataframe
    df_list[[i]] <- data.frame(
      #rgn = i,
    area = paste(round(area_per_eez / 1e8, 2), "x 10^8")
    )
  }

# The dataframe
df_suit_area_eez <- do.call(rbind, df_list)
  
# Output a kable table
kable(
    df_suit_area_eez,
    align = 'c',
    col.names = c("EEZ location", "Suitable Habitat Area")
  )
  
```

Oysters have a large amount of suitable habitats in Washington, Central CA, and Southern CA. Central and South California's temperature, ranging from 12°C to 20°C, is fully within the limits of oysters preferences. Additionally, oysters have a depth span of 0-70 meters below sea level, meaning oysters will reside closer to the coastline. The Washington and South California's Exclusive Economic Zones offer a larger continental shelf where oysters can thrive. In conclusion, when mixing oysters depth and temperature preferences, Washington, Central CA, and Southern CA offer the most suitable habitats.

------------------------------------------------------------------------

## The Function

Create a function that does everything above for you

```{r}

habitat <- function(depth_df,
                    temp_df,
                    depth_min,
                    depth_max,
                    temp_min,
                    temp_max,
                    species_name,
                    eez_df) {
  
  # PART 1: MATRIX, RECLASSIFY, COMBINED, MASK, TOTAL AREA ********************************** 
  # Reclass matrix - DEPTH
  rcl_depth <- matrix(
    c(-Inf, depth_min, 0, # SUITABLE depth
      depth_min, depth_max, 1, #  SUITABLE depth
      depth_max, Inf, 0), # SUITABLE depth
    ncol = 3,
    byrow = TRUE
  )
  # Reclass matrix - TEMP
  rcl_temp <- matrix(
    c(-Inf, temp_min, 0, # UNSUITABLE temp
      temp_min, temp_max, 1, #SUITABLE temp
      temp_max, Inf, 0), #UNSUITABLE temp
    ncol = 3,
    byrow = TRUE
  )
  
  
  # Apply the reclassification matrix to the raster
  depth_reclass <- classify(depth_df, rcl_depth)
  temp_reclass <- classify(temp_df, rcl_temp)
  
  # Check (and change crs if necessary)
  if (st_crs(temp_reclass) == st_crs(depth_reclass)) {
    print("CRS match")
  } else {
    warning("Updating CRS")
    temp_reclass <- st_transform(temp_reclass, crs = crs(depth_reclass))
  }
  
  # Multiply rasters
  suit_raster <- temp_reclass * depth_reclass
  
  # Crop to eez area
  species_raster_pre <- mask(suit_raster, eez_df)
  
  # Select only suitable habitats
  species_raster <- species_raster_pre
  species_raster[species_raster  == 0] <- NA
  
  # Area of all suitable habitat
  area <- expanse(species_raster)[1, 2]
  
  print(paste(
      "The area of all suitable",
      species_name,
      "habitats in the West Coast EEZs is",
      round(area / 1e9, 2),
      "x 10^9 m^2"
    )
  )
  
   # Check (and change crs if necessary)
  if (crs(species_raster) == crs(eez_df)) {
    print("CRS match")
  } else {
    warning("Updating CRS")
    eez_df <- transform(eez_df, crs = crs(species_raster))
  }

  # PART 2: FIND AREA PER EEZ --> DATAFRAME ********************************* 
  # Empty list to feed for loop into after
  df_list <- list()
  
  # For loop to get area of each eez
  for (i in unique(eez_df$rgn)) {
    area_eez <- eez_df %>% filter(rgn == i) #
    spec_per_eez <- mask(species_raster, area_eez)
    area_per_eez <- expanse(spec_per_eez)[1, 2]
    
    # Create dataframe
    df_list[[i]] <- data.frame(
      #rgn = i,
      area = area_per_eez
      #paste(round(area_per_eez / 1e8, 2), "x 10^8")
    )
    
  }
  
  
  # The dataframe for the table 
  df_suit_area_eez <- do.call(rbind, df_list)
  # geodata frame for the map 
  eez_df[, 'area_suit_hab'] <- NA # Create new column 
  eez_df$area_suit_hab <- df_suit_area_eez$area
  
  
  # Output a kable table
  print(kable(
    df_suit_area_eez,
    align = 'c',
    format.args = list(scientific = TRUE, digits = 3), 
    col.names = c("EEZ location", "Suitable Habitat Area (m)")
   ))
  
  
  # printing geodataframe 
  #print(eez_df)
  
  
  
  # PART 2: MAKING MAPS ********************************* 
  
  # DEPTH AND TEMP MAPS
  
  # Plotting suitable habitats for Oysters
  map_temp <- tm_shape(temp_reclass) +
    tm_raster(
      col.scale = tm_scale_categorical(
        labels = c("Unsuitable (0)", "Suitable (1)"),
        values = c('lightblue2', 'steelblue4'),
      ),
      col.legend = tm_legend(title = "Suitable vs Unsuitable Temp")
    ) +
    tm_title(text = paste("Suitable Temp for", species_name)) +
    tm_compass(position = c('right', 'top'))
  
  
  map_depth <- tm_shape(depth_reclass) +
    tm_raster(
      col.scale = tm_scale_categorical(
        labels = c("Unsuitable (0)", "Suitable (1)"),
        values = c('lightblue2', 'steelblue4'),
      ),
      col.legend = tm_legend(title = paste("Suitable vs Unsuitable Depth"))
    ) +
    tm_title(text = paste("Suitable Depth for", species_name))
  
  
  
  map_both <- tm_shape(suit_raster) +
    tm_raster(
      col.scale = tm_scale_categorical(
        labels = c("Unsuitable (0)", "Suitable (1)"),
        values = c('lightblue2', 'red'),
      ),
      col.legend = tm_legend(title = "Suitable vs Unsuitable Habitat")
    ) +
    tm_shape(eez_df) + 
      tm_polygons(
        fill_alpha = 0, 
        col = 'black', 
        lwd = 0.2) + 
    tm_title(text = paste("Suitable Depth & Temp for", species_name)) +
    tm_scalebar(position = c('left', 'bottom'))
  
  print(tmap_arrange(map_temp, map_depth, map_both))
  
  
  
  # MAIN MAP OF SUITABLE HABITAT
  output_map <- tm_shape(eez_df, 
                         bbox = sf::st_bbox(c(
                           xmin = -131, xmax = -110,
                           ymin = 30,  ymax = 50))) +
    tm_polygons(
      fill = 'area_suit_hab', 
      fill.legend = tm_legend(title = "Total Suitable\nHabitat Area (km^2)", 
                              # orientation = 'landscape', 
                              # position = tm_pos_out("center", "bottom", pos.h = "center")
                              ),
      #fill.scale = tm_scale(values = 'Brewer.yl_gn') 
      fill.scale = tm_scale_continuous(n = 4, 
                                       values = 'brewer.YlOrBr'), 
    ) +
    tm_title(text = paste("Suitable Habitat for", species_name, "within\nthe West Coast Exclusive Economic Zones")) +
    tm_scalebar(position = c('right', 'top'), margins = 1) +
    tm_compass(position = c('right', 'bottom')) +
    tm_basemap("Esri.OceanBasemap") + 
    tm_text('rgn', size = 0.5) + 
    tm_credits(text = paste0(species_name," perfer depths of ", depth_min," to ", depth_max, "(m) and temps of ", temp_min, " to ", temp_max, "°C.")) + 
     tm_layout(legend.outside.position = "bottom")

  
  print(output_map)
  
  
}


```

Apply the Function to Oysters - Sea surface temperature: 11-30°C - Depth: 0-70 meters below sea level

```{r}

# Use function 
df_suit_area_eez <- habitat(
    depth_df = bathy_updated,
    temp_df = sst_mean, 
    depth_min = -70,
    depth_max = 0,
    temp_min = 11,
    temp_max = 30,
    species_name = 'Oysters',
    eez_df = eez
) 

```

------------------------------------------------------------------------

## Suitable locations for [Siliqua-patula]{.underline}

The Pacific razor clam location ranges from the Aleutian Islands in Alaska to Pismo Beach, California. The clam is typically discovered in depths between 0-55m, and its preferred temperature is around 5°C (estimating 0°C - 10°C).

Therefore, Pacific razor clam research has shown that the clams need the following conditions for optimal growth:

-   Sea surface temperature: 0-10°C

-   Depth: 0-55 meters below sea level

### Use Function to Find Suitable Habitat for clams in West Coast EEZs

```{r}

# Use function 
df_suit_area_eez <- habitat(
    depth_df = bathy_updated,
    temp_df = sst_mean, 
    depth_min = 0,
    depth_max = 55,
    temp_min = 0,
    temp_max = 10,
    species_name = 'Clams',
    eez_df = eez
) 

```

### Additional Map For Clams

I wanted to see Washington better, so I created tailored map to look at Washington

```{r}
#| code-fold: true
#| warning: false

# Creating dataframes in order to plot 
rcl_depth_clam <- matrix(c(-Inf, -55, 0, # UNSUITABLE
                      -55, 0, 1, #  SUITABLE HABITAT
                      0, Inf, 0), # UNSUITABL
                      ncol = 3, byrow = TRUE)
# Temp matrix 
rcl_temp_clam <- matrix(c(-Inf, 0, 0, # UNSUITABLE
                      0, 10, 1, #  SUITABLE HABITAT
                      10, Inf, 0), # UNSUITABL
                      ncol = 3, byrow = TRUE)

# Apply the reclassification matrix to the depth and temp rasters 
depth_reclass_clam <- classify(bathy_updated, rcl_depth_clam)
temp_reclass_clam <- classify(sst_mean, rcl_temp_clam)

# Remove matrixs as objects 
rm(list = c('rcl_depth_clam', 'rcl_temp_clam'))

# Multiple temp and depth rasters 
clam_suit_raster <- depth_reclass_clam * temp_reclass_clam
# Crop to EEZs 
clam_suit_crop <- mask(clam_suit_raster, eez) 



# bbox of Washington EEZ 
washington <- eez %>% filter(rgn_key == 'WA')
clam_suit_washington <- mask(clam_suit_crop, washington)



# Inset map of Just clams in Washington 
washington_inset <- tm_shape(
   clam_suit_washington, 
   bbox = sf::st_bbox(c(
    xmin = -129.5, xmax = -122,
    ymin = 45.5,  ymax = 49
  )), 
  crs = st_crs(clam_suit_washington) ) + 
  tm_raster( 
    col.scale = tm_scale_categorical(
      values = c('lightblue2', 'red')), 
    col.legend = tm_legend_hide()) +
  tm_shape(washington) + 
  tm_polygons(lwd = 0.35, 
              fill = "peachpuff2", 
              fill_alpha = 0.2) +
  tm_title(text = "Washington's EEZ", 
           position = tm_pos_in(), 
           size = 0.8) +
 tm_basemap("Esri.OceanBasemap") + 
  tm_scalebar(position = c('right', 'bottom'))

 
# TEMP AND DEPTH AND EEZ MAP 
tm_shape(
  clam_suit_crop, 
     bbox = sf::st_bbox(c(
    xmin = -131, xmax = -110,
    ymin = 30,  ymax = 50
  )), 
  crs = st_crs(clam_suit_crop)) + 
  tm_raster(
    col.scale = tm_scale_categorical(
      labels = c("Unsuitable (0)", "Suitable (1)"), 
      values = c('lightblue2', 'red'),
    ),
    col.legend = tm_legend(title = "Suitable vs Unsuitable\nOysters Habitat", 
                           position = tm_pos_in('right', 'center'), 
                           title.size = 0.6, 
                           text.size = 0.5, 
                           bg.alpha = 0.5)
    ) + 
  tm_title(text = "Suitable Depth & Temp for Clams\nwithin the West Coast EEZs") + 
  tm_shape(eez) + 
  tm_polygons(lwd = 0.35, 
              fill = "peachpuff2", 
              fill_alpha = 0.2) + 
  tm_basemap("Esri.OceanBasemap") +
  tm_compass(position = c('left', 'bottom'))+ 
  tm_scalebar(position = c('left', 'bottom')) +
  tm_inset(washington_inset, 
           position = tm_pos_out(), 
           width = 15, 
           height = 15)



# Remove objects 
rm(list = c('map_temp_clam', 'map_depth_clam', 'map_clam', 'clam_suit_washington', 'washington'))


```


The Pacific razor clam requires cold temperatures ranging from  0  to 10°C and shallow depths from 0 to 55m, This pins the clams to the coastline and to the North. Hence, Washington's Exclusive Economic Zone has the most suitable habitat for the Pacific razor clam. 



## Reflection 

The suitable habitat function creates a reclassification matrix based on the depth and temperature parameters. The reclassification matrix is then turned into a raster of the species' suitable habitat, which is then cropped and masked within the West coast exclusive economic zones (EEZs). Subsequently, the function filters through and finds the area of suitable habitat for each EEZ (Washington, Oregon, North, Central, and South California). Finally, the function creates a table and multiple maps to display suitable depth, temperature, the interaction of the two, and the area per each EEZ. 

Within this study, suitable habitat is only dictated by depth and temperature ranges. However, habitat suitability includes other features, such as water quality, human disturbance, and physical structures. Therefore, the lack of other important parameters limits this study. Furthermore, the function used the average temperature and depth measurements, not accounting for climate variation. In future studies, other parameters such as seasonality and water quality data could be added into the function to improve accuracy. 


## References 
Hall, S. J., Delaporte, A., Phillips, M. J., Beveridge, M. & O’Keefe, M. Blue Frontiers: Managing the Environmental Costs of Aquaculture (The WorldFish Center, Penang, Malaysia, 2011).↩︎

Gentry, R. R., Froehlich, H. E., Grimm, D., Kareiva, P., Parke, M., Rust, M., Gaines, S. D., & Halpern, B. S. Mapping the global potential for marine aquaculture. Nature Ecology & Evolution, 1, 1317-1324 (2017).↩︎

GEBCO Compilation Group (2022) GEBCO_2022 Grid (doi:10.5285/e0f0bb80-ab44-2739-e053-6c86abc0289c)